### 2013-12-12 doc/Review1Response.txt
Addressed reviewer ones comments, relagating a lot of the information to appendecies

1st referee

The paper presents a numerical calculation on the measurement of BNS physical parameters, from GW signals to be obtained with the advanced detector array.
The paper is clearly written and the calculation look sound and free of basic errors.
The analysis reported here is a sort of reduced version, at higher SNR, of the analysis reported in Nissanke et al, 2013, Astrophys. J., 767, 124, 124.
Both analyses use the zero-noise, MCMC technique, and their results are consistent.
I am not convinced that the paper, at this stage of the work, provides substantial new information over that of Nissanke. The only piece of information that I find new is to be found in figure 7 and 8, wherein the positional accuracy is reported as a function of the sky location of the source.
As most of methods and procedures are not new, I think the authors, after having substantially shortened the introductory parts and minimized the description of the zero-noise MCMC method, may want to highlight the delta knowledge contributed by this paper relative to the recent literature.

I have also one comment on section 3.4 and the related figure 9.

At face value figure 9 would indicate that by averaging overn many non zero-noise MCMC calculation, the distribution of all parameters would be substantially smeared out . The authors state instead that the figure demonstrates that the distribution stays the same as in the zero-noise case. I think this is confusing and should be clarified.

I think the authors have to respond to these objections before the paper may be considered for publication.

———————————————————————————————————

The reviewer brings up two points where our paper can be improved: the length of the description of our MCMC zero-noise choice, and the difference between our work and the work of Nissanke et al.  We discuss these points below:

Length of the zero-noise and introductory sections:
	We completely agree with the reviewer that these sections can be substantially paired down.  We have made extensive modifications to sections 2.1 and 2.2, relegating the majority of the technical details in those sections to Appendecies.  We find this improves  the flow of Section two.  Additionally, we have moved the majority of the discussion regarding the zero-noise choice (the end of section 2.3 and all of section 3.4) to a third Appendix.

Distinction between the current study and Nissanke et al. 2013:

The second issue is, of course, what additional information is provided by our paper that was not provided by Nissanke at al. 2013; however, we believe that our study is significantly different from Nissanke2013.  The numbers in Tables 1 and 2, which provide a “rule-of-thumb” estimate for the community, were in many ways the main goal of this paper. As Nissanke2013 primarily reports on EM follow-up capabilities, we feel that the current study is significantly different.

The current study is the first to examine the intrinsic properties of neutron star binaries, in particular the masses.  Our current study focuses heavily on the recovery of the mass parameters for multiple neutron star masses.  This is reported in our section 3.1, and the tables in section 4.  Nissanke2013 does not consider the recovery of the BNS mass parameters, and only simulates neutron stars of 1.4 solar masses, as their study is primarily interested in extrinsic parameters relevant to electromagnetic followup.

We also feel that our section 3.3, which concerns the sky-localization, is distinct from Nissanke2013.  As the reviewer points out, our figures 6 and 7 demonstrate the uncertainties over a large number of sky-locations for each detector network, while Nissanke2013 only demonstrates a handful of injections.  In fact, the realistic population plotted in Figure 8 is taken directly from Nissanke2013.

Finally, the reviewer points out that the caption for Figure 9 is confusing.  We agree, and have clarified the text (specifically mentioning that the variance and mean of multiple zero-noise results are accurate to different orders in 1/SNR)